### 2012-06-13 inst/examples/Reed.Rmd
Reed using new extension, plotting policy function

`ro cache=TRUE, tidy=FALSE, warning=FALSE, comment=NA, message=FALSE, refresh=2 or`

``` {r  echo=FALSE, cache=FALSE }
render_jekyll()
opts_knit$set(upload.fun = socialR::flickr.url)
options(device = function(width = 5, height = 5) {
    pdf(NULL, width = width, height = height)
})
knitcitations::cleanbib()
````

# Sethi Model
 * author Carl Boettiger, <cboettig@gmail.com>
 * license: CC0

 Implements a numerical version of the SDP described in `r knitcitations::citep("10.1016/j.jeem.2004.11.005")`.

Clear the workspace and load package dependencies: 
``` {r  setup, echo=FALSE}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
rm(list=ls())
````

Define noise parameters

``` {r }
sigma_g <- 0.1    # Noise in population growth
sigma_m <- 0.1     # noise in stock assessment measurement
sigma_i <- 0.1     # noise in implementation of the quota
````

we'll use log normal noise functions. 
For Reed, only `z_g` will be random.
Sethi et al will add the other forms of noise:

@@

That is, \\( f(x,h) = \frac{A x}{1 + B x} \\)

@@

We must now define parameters for the function.  Note that the positive stationary root of the model is given by \\( \frac{A-1}{B} \\), which we'll store for future reference as `K`.

@@

``` {r  profit}
profit <- profit_harvest(price=1, c0 = 0.01) 
````

The `profit_harvest` function has the form \\( \Pi = h - \left( c_0  + c_1 \frac{h}{x} \right) \frac{h}{x} \\), conditioned on \\( h > x \\) and \\(x > 0 \\).  Note that the R code defines a function from another function using a trick known as a _closure_.  Again we could write a custom profit function as long as it can take a vector stock size `x` and a scalar harvest level `h`.  Details for provided functions can be found in the manual, i.e. `?profit_harvest`.

@@

### Calculate the transition matrix (with noise in growth only)

We calculate the stochastic transition matrix for the probability of going from any state \\(x_t \\) to any other state \\(x_{t+1}\\) the following year, for each possible choice of harvest \\( h_t \\).  This provides a look-up table for the dynamic programming calculations.

In the Sethi case, computing the distribution over multiple sources of noise is actually quite difficult.  Simulation turns out to be more efficient than numerically integrating over each distribution.  This code parallelizes the operation over four cores, but can be scaled to an arbitrary cluster.

``` {r  parallel}
require(snowfall) 
sfInit(parallel=TRUE, cpu=4)
SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
````
Note that `SDP_Mat` is specified from the calculation above, as are our grids and our profit function. `OptTime` is the stopping time.  `xT` specifies a boundary condition at the stopping time. A reward for meeting this boundary must be specified for it to make any difference.  `delta` indicates the economic discount rate. Again, details are in the function documentation.

@@

### Simulate

Now we'll simulate 100 replicates of this stochastic process under the optimal harvest policy determined above.

@@

The forward simulation algorithm needs an initial condition `x0` which we set equal to the carrying capacity, as well as our population dynamics `f`, parameters `pars`, grids, and noise coefficients.  Recall in the Reed case only `z_g`, growth, is stochastic, while in the Sethi example we now have three forms of stochasticity -- growth, measurement, and implementation.

@@

``` {r  onerep }
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, fishstock)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_line(aes(time, harvest), col="darkgreen") 
````

This plot summarizes the stock dynamics by visualizing the replicates. Reed's S shown again.

``` {r  all}
p1 <- ggplot(dt) + geom_abline(intercept=opt$S, slope = 0) 
p1 + geom_line(aes(time, fishstock, group = reps), alpha = 0.2)
````

@@

``` {r  harvestplot}
p1 + geom_line(aes(time, harvest, group = reps), alpha = 0.1, col="darkgreen")
````

@@

``` {r  escapement}
p1 + geom_line(aes(time, escapement, group = reps), alpha = 0.1, col="darkgrey")
````

### Visualizing the optimal policy

Note that when the boundary is sufficiently far away, i.e. for the first couple timesteps, the optimal policy is stationary.  The optimal policy is shown here over time, where the color indicates the harvest recommended for each possible stock value at that time (shown on the vertical axis). Observe that the solution does not correspond to a simple Reed-style rule.

``` {r  policy, fig.width=9}
policy <- melt(opt$D)
policy_zoom <- subset(policy, x_grid[Var1] < max(dt$fishstock) )
p5 <- ggplot(policy_zoom) + 
  geom_point(aes(Var2, (x_grid[Var1]), col=h_grid[value])) + 
  labs(x = "time", y = "fishstock") +
  scale_colour_gradientn(colours = rainbow(4)) +
  geom_abline(intercept=opt$S, slope = 0) 
p5
````

@@

``` {r  policy2, fig.width=10}
p6 <- ggplot(policy_zoom) + 
  geom_point(aes(Var2, (x_grid[Var1]), col=x_grid[Var1] - h_grid[value])) + 
  labs(x = "time", y = "fishstock") +
  scale_colour_gradientn(colours = rainbow(4)) +
  geom_abline(intercept=opt$S, slope = 0) 
p6 + geom_line(aes(time, fishstock, group = reps), alpha = 0.1, data=dt)
````

# References

``` {r biblio, echo=FALSE, results="asis"}
I(knitcitations::bibliography())
````

>>>>

`ro cache=FALSE, tidy=FALSE, warning=FALSE, comment=NA, message=FALSE, verbose=TRUE or`

``` {r  echo=FALSE }
opts_knit$set(upload.fun = socialR::flickr.url)
options(device = function(width = 5, height = 5) {
    pdf(NULL, width = width, height = height)
})
````

# Reed Model
 * author Carl Boettiger, <cboettig@gmail.com>
 * license: CC0

 Implements a numerical version of the SDP described in:

   Reed, W.J., 1979. Optimal Escapement Levels in Stochastic
   and Deterministic Harvesting Models. Journal of Environmental 
   Economics and Management. 6: 350-363.

Clear the workspace and load package dependencies: 
``` {r  setup, echo=FALSE}
rm(list=ls())   
require(pdgControl)
require(reshape2)
require(ggplot2)
require(data.table)
````

### Define parameters

``` {r }
gridsize <- 100   # gridsize (discretized population)
sigma_g <- 0.2    # Noise in population growth
sigma_m <- 0.     # noise in stock assessment measurement
sigma_i <- 0.     # noise in implementation of the quota
````

we'll use log normal noise functions. 
For Reed, only `z_g` will be random.
Sethi et al will add the other forms

@@

That is, \( f(x,h) = \frac{A x}{1 + B x} \)

@@

We must now define parameters for the function.  Note that the positive stationary root of the model is given by \( (A-1)/B \), which we'll store for future reference as `K`.

@@

``` {r  profit}
profit <- profit_harvest(price=1, c0 = 0.1) 
````

The profit_harvest function has the form \( \Pi = h - \left( c_0  + c_1 \frac{h}{x} \right) \frac{h}{x} \), conditioned on \( h > x \) and \(x > 0 \).  Note that the R code defines a function from another function using a trick known as a _closure_.  Again we could write a custom profit function as long as it can take a vector stock size `x` and a scalar harvest level `h`.  Details for provided functions can be found in the manual, i.e. `?profit_harvest`.

@@

### Calculate the transition matrix (with noise in growth only)      
We calculate the stochastic transition matrix for the probability of going from any state \(x_t \) to any other state \(x_{t+1}\) the following year, for each possible choice of harvest \( h_t \).  This provides a look-up table for the dynamic programming calculations.  Take a look at the R code and the documentation for `determine_SDP_matrix` to see what this function is actually doing.  The implementation is quite simple.

``` {r  SDP_Mat}
SDP_Mat <- determine_SDP_matrix(f, pars, x_grid, h_grid, sigma_g )
````

@@

Note that `SDP_Mat` is specified from the calculation above, as are our grids and our profit function. `OptTime` is the stopping time.  `xT` specifies a boundary condition at the stopping time. A reward for meeting this boundary must be specified for it to make any difference.  `delta` indicates the economic discount rate. Again, details are in the function documentation.

We can take a look at the policy function,

``` {r }
qplot(x_grid, opt$D[,1], xlab="stock size", ylab="harvest")
````

In the Sethi case, computing the distribution over multiple sources of noise is actually quite difficult.  Simulation turns out to be more efficient than numerically integrating over each distribution.  This code parallelizes the operation over four cores, but can be scaled to an arbitrary cluster. Since we're focused on the Reed example for the moment, we can ignore this step.

``` {r  parallel}
require(snowfall) 
sfInit(parallel=TRUE, cpu=4)
sethi_SDP_Mat <- SDP_by_simulation(f, pars, x_grid, h_grid, z_g, z_m, z_i, reps=999)
sethi_opt <- find_dp_optim(sethi_SDP_Mat, x_grid, h_grid, OptTime=25, xT=0, 
                     profit, delta=0.05, reward=0)
````

We can take a look at the stochastically estimated policy function,

``` {r }
qplot(x_grid, sethi_opt$D[,1], xlab="stock size", ylab="harvest")
````

### Simulate 
Now we'll simulate 100 replicates of this stochastic process under the Reed optimal harvest policy determined above.

@@

The forward simulation algorithm needs an initial condition `x0` which we set equal to the carrying capacity, as well as our population dynamics `f`, parameters `pars`, grids, and noise coefficients.  Recall in the Reed case only `z_g`, growth, is stochastic.

@@

``` {r  p0}
ggplot(subset(dt,reps==1)) +
  geom_line(aes(time, fishstock)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_line(aes(time, harvest), col="darkgreen") 
````

This plot summarizes the stock dynamics by visualizing the replicates. Reed's S shown again, along with the dotted line showing the allee threshold, below which the stock will go to zero (unless rescued stochastically).

``` {r  p1}
p1 <- ggplot(dt) + geom_abline(intercept=opt$S, slope = 0) + 
  geom_abline(intercept=xT, slope = 0, lty=2) 
p1 + geom_line(aes(time, fishstock, group = reps), alpha = 0.2)
````

@@

``` {r  p2}
p1 + geom_line(aes(time, harvest, group = reps), alpha = 0.1, col="darkgreen")
````

@@

``` {r  p3}
p1 + geom_line(aes(time, escapement, group = reps), alpha = 0.1, col="darkgrey")
````

### Visualizing the optimal policy
Note that when the boundary is sufficiently far away, i.e. for the first couple timesteps, the optimal policy is stationary.  The optimal policy is shown here over time, where the color indicates the harvest recommended for each possible stock value at that time (shown on the vertical axis).  Note that below a certain stock value, harvesting is not recommended and the dots turn red (Reed's constant escapement rule!)  However, at very low values, harvesting starts again (orange dots), because of the allee effect - these populations are doomed anyway, so may as well fish all that remains.

Note that interestingly, populations just below the allee threshold are given the chance to be rescued stochastically early on - that small chance that they recover is worth the expected loss.  The "no-harvest" zones stand out clearly in the red areas of this graph.

``` {r  policy, fig.width=9}
policy <- melt(opt$D)
policy_zoom <- subset(policy, x_grid[Var1] < max(dt$fishstock) )
p5 <- ggplot(policy_zoom) + 
  geom_point(aes(Var2, (x_grid[Var1]), col=h_grid[value])) + 
  labs(x = "time", y = "fishstock") +
  scale_colour_gradientn(colours = rainbow(4)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_abline(intercept=xT, slope=0, lty=2)
p5
````

@@

``` {r  policy2, fig.width=10}
p6 <- ggplot(policy_zoom) + 
  geom_point(aes(Var2, (x_grid[Var1]), col=x_grid[Var1] - h_grid[value])) + 
  labs(x = "time", y = "fishstock") +
  scale_colour_gradientn(colours = rainbow(4)) +
  geom_abline(intercept=opt$S, slope = 0) +
  geom_abline(intercept=xT, slope=0, lty=2)
p6 + geom_line(aes(time, fishstock, group = reps), alpha = 0.1, data=dt)
````
